---
title: "UCSC RepeatMasker AnnotationHub Resource Metadata"
author:
- name: Robert Castelo
  affiliation: 
  - &id Dept. of Experimental and Health Sciences, Universitat Pompeu Fabra, Barcelona, Spain
  email: robert.castelo@upf.edu
package: "`r pkg_ver('UCSCRepeatMasker')`"
abstract: >
  Store UCSC RepeatMasker AnnotationHub resource metadata.
vignette: >
  %\VignetteIndexEntry{UCSC RepeatMasker AnnotationHub Resource Metadata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, echo=FALSE}
options(width=80)
```

# Retrieval of UCSC RepeatMasker annotations through AnnotationHub resources

The `UCSCRepeatMasker` package provides metadata for the
`r Biocpkg("AnnotationHub")` resources associated with UCSC RepeatMasker
annotations. The original data can be found through UCSC download URLs
`https://hgdownload.soe.ucsc.edu/goldenPath/XXXX/database/rmsk.txt.gz`,
where `XXXX` is the corresponding code to a UCSC genome version.
Details about how those original data were processed into
`r Biocpkg("AnnotationHub")` resources can be found in the source
file:

```
UCSCRepeatMasker/scripts/make-metadata_UCSCRepeatMasker.R
```

The UCSC RepeatMasker annotations can be retrieved using the
`r Biocpkg("AnnotationHub")`,
which is a web resource that provides a central location where genomic files
(e.g., VCF, bed, wig) and other resources from standard (e.g., UCSC, Ensembl)
and distributed sites, can be found. A Bioconductor `r Biocpkg("AnnotationHub")`
web resource creates and manages a local cache of files retrieved by the user,
helping with quick and reproducible access.

For example, to list the available UCSC RepeatMasker annotations for the human
genome we should query the `r Biocpkg("AnnotationHub")` as follows:

```{r message=FALSE, cache=FALSE, eval=FALSE}
library(AnnotationHub)

ah <- AnnotationHub()
query(ah, c("RepeatMasker", "Homo sapiens"))
```

We finally retrieve the desired resource with the following notation:

```{r message=FALSE, cache=FALSE, eval=FALSE}
rmskhg38 <- ah[["AHXXXX"]]
```
where `AHXXXX` is the identifier of the desired resource.


# Session information

```{r session_info, cache=FALSE}
sessionInfo()
```
